Stateless Model Checking with Data-Race Preemption Points

### 2015-11-16 pldi-16/overview.tex
Fix pictures

\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& \multicolumn{2}{c}{\texttt{int x = 0; mutex\_t mx;}} \\
	& {\bf Thread 1} & {\bf Thread 2} \\
	1 & \texttt{\hilight{orange}{mutex\_lock(\&mx);}} & \\
	2 & \texttt{int tmp = x;} &\\
	3 & & \texttt{atomic\_xadd(\&x, 1);} \\
	4 & & \texttt{\hilight{olivegreen}{yield();}} \\
	5 & & \texttt{atomic\_xadd(\&x, 1);} \\
	6 & \texttt{x = tmp + 1;} & \\
	7 & \texttt{\hilight{olivegreen}{yield();}} & \\
	8 & \texttt{\hilight{commentblue}{mutex\_unlock(\&mx);}} & \\
	9 & \texttt{assert(x >= 1);}  & \\
	10 & & \texttt{assert(x >= 2);} \\
	%1 & \texttt{\hilight{brickred}{x->foo = ...;}} & \\
	%2 & \texttt{\hilight{olivegreen}{free}(x);} \\
	%3 & & \texttt{\hilight{commentblue}{// x's memory recycled}} \\
	%4 & & \texttt{y~=~\hilight{olivegreen}{malloc}(sizeof *y);} \\
	%5 & & \texttt{\hilight{commentblue}{// ...initialize...}}\\
	%6 & & \texttt{publish(y);} \\
	%7 & & \texttt{\hilight{brickred}{y->bar = ...;}} \\
\end{tabular}
\caption{Example program with a data-race bug. In this interleaving, the assertion on line 10 will fail. Two data-race preemptions are required to expose the bug: one just before thread 1's line 6, and one just before thread 2's line 10.}
\label{fig:example}
\end{figure}

@@

% TODO: Fix picture.
\begin{figure}[t]
	\includegraphics[width=0.48\textwidth]{trees.pdf}
	\caption{Iterative Deepening example.
		The minimal state space (top) includes only voluntary thread switches, such as {\tt yield()}. %or {\tt cond\_wait()}.
		Multiple further tests can be run: preempting on calls to {\tt mutex\_lock} alone (left), {\tt mutex\_unlock} alone (right), or both together (bottom).
Each option increases the state space size unpredictably, so multiple state spaces should be tested in parallel.
Estimation techniques~\cite{estimation} inform which state spaces to prioritize.
}
	\label{fig:id}
\end{figure}

The checker defins the granularity of thread interleavings by which {\em preemption points} it uses to switch threads.
Most model checkers \cite{chess} choose synchronization/thread library API boundaries for these points;
in our example program, these would be lines 1, 4, 7, and 8.
Figure~\ref{fig:id} shows several possible resulting state spaces.
The approach of prior work is to enable all preemption points simultaneously, i.e., to test only the bottom state space.

@@

Data race detection \cite{eraser} is an analysis which identifies pairs of unsynchronized memory accesses between threads.
Two instructions are said to race if
they both access the same memory address,
at least one is a write,
the threads do not hold the same lock,
and no synchronization enforces the order of the thread transitions.
In Figure~\ref{fig:example}, lines 3 and 5 each race with 2, 6, and 9, and line 6 further races with 10.

Though state-of-the-art MCs preempt only on synchronization events,
many serious concurrency bugs are caused by data races leading to corrupted shared state.
Figure~\ref{fig:example}'s thread interleaving is possible only with {\em data-race preemption points}:
preempting a thread just before an instruction identified as part of a data race.
None of the state spaces in Figure~\ref{fig:id} contain this interleaving.

>>>>

\begin{figure}[t]
	\small
\begin{tabular}{rll}
	& \multicolumn{2}{c}{\texttt{int x = 0; mutex\_t mx;}} \\
	& {\bf Thread 1} & {\bf Thread 2} \\
	1 & \texttt{\hilight{orange}{mutex\_lock(\&mx);}} & \\
	2 & \texttt{int tmp = x;} &\\
	3 & & \texttt{atomic\_xadd(\&x, 1);} \\
	4 & & \texttt{\hilight{olivegreen}{yield();}} \\
	5 & & \texttt{atomic\_xadd(\&x, 1);} \\
	6 & \texttt{x = tmp + 1;} & \\
	7 & \texttt{\hilight{commentblue}{mutex\_unlock(\&mx);}} & \\
	8 & \texttt{assert(x >= 1);}  & \\
	9 & & \texttt{assert(x >= 2);} \\
	%1 & \texttt{\hilight{brickred}{x->foo = ...;}} & \\
	%2 & \texttt{\hilight{olivegreen}{free}(x);} \\
	%3 & & \texttt{\hilight{commentblue}{// x's memory recycled}} \\
	%4 & & \texttt{y~=~\hilight{olivegreen}{malloc}(sizeof *y);} \\
	%5 & & \texttt{\hilight{commentblue}{// ...initialize...}}\\
	%6 & & \texttt{publish(y);} \\
	%7 & & \texttt{\hilight{brickred}{y->bar = ...;}} \\
\end{tabular}
\caption{Example program with a data-race bug. In this interleaving, the assertion on line 9 will fail. Two data-race preemptions are required to expose the bug: one just before thread 1's line 6, and one just before thread 2's line 9.}
\label{fig:example}
\end{figure}

@@

\begin{figure}[t]
	\includegraphics[width=0.48\textwidth]{trees-v2.pdf}
	\caption{Iterative Deepening example.
		The minimal state space (top) includes only voluntary thread switches, such as {\tt yield()}. %or {\tt cond\_wait()}.
		Multiple further tests can be run: preempting on calls to {\tt mutex\_lock} alone (left), {\tt mutex\_unlock} alone (right), or both together (bottom).
Each option increases the state space size unpredictably, so multiple state spaces should be tested in parallel.
Estimation techniques~\cite{estimation} inform which state spaces to prioritize.
}
	\label{fig:id}
\end{figure}

The checker defins the granularity of thread interleavings by which {\em preemption points} it uses to switch threads.
Most model checkers \cite{chess} choose synchronization/thread library API boundaries for these points;
in our example program, these would be lines 1, 4, and 7.
Figure~\ref{fig:id} shows several possible resulting state spaces.
The approach of prior work is to enable all preemption points simultaneously, i.e., to test only the bottom state space.

@@

Data race detection \cite{eraser} is an analysis which identifies pairs of unsynchronized memory accesses between threads.
Two instructions are said to race if
they both access the same memory address,
at least one is a write,
the threads do not hold the same lock,
and no synchronization enforces the order of the thread transitions.
In Figure~\ref{fig:example}, lines 3 and 5 each race with 2, 6, and 8, and line 6 further races with 9.

Though state-of-the-art MCs preempt only on synchronization events,
many serious concurrency bugs are caused by data races leading to corrupted shared state.
Figure~\ref{fig:example}'s thread interleaving is possible only with {\em data-race preemption points}:
preempting a thread just before an instruction identified as part of a data race.
None of the state spaces in Figure~\ref{fig:id} contain this interleaving,
as none of the mutex/yield preemptions split lines 2 and 6 across different transitions.
